Fundamental Problems with the Evidence Base of Adolescent Depression Treatments

### 2024-03-25 Untitled.qmd
Fixing up tables so that they render to word - using flextable

---
title: "Fundamental Problems with the Evidence Base of Adolescent Depression Treatments"
author: "Stringaris et al"
prefer-html: true
format: docx
editor: visual
---

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
```

@@

df_long_for_metan <- df_long_for_metan %>%
  rowwise() %>%
  mutate(responders = sum(rnorm(baseline_n, post_mean, 
                                post_sd) < (baseline_mean/ 2)))

df_long_for_metan <- df_long_for_metan %>%
  rowwise() %>%
  mutate(resp_rate = responders/baseline_n)

summary_resp_rate <- df_long_for_metan %>%
  group_by(group) %>%
  summarise(mean_resp_rate = mean(resp_rate, na.rm = TRUE),
            sd_resp_rate = sd(resp_rate, na.rm = TRUE))

test_summary <- df_with_distinct_instruments %>% 
  group_by(psy_or_med) %>% 
  summarise(rr_active = mean(resp_rate_active, na.rm = TRUE),
            rr_control = mean(resp_rate_control, na.rm = TRUE),
            n = n())

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Now start looking at age variable

@@

# Rounding
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3))

colnames(df_overall_baseline_results)[8] <- paste0(colnames(df_overall_baseline_results)[8], footnote_marker_number(1))

#df_overall_baseline_results <- options(knitr.kable.NA = "") # CHARLOTTE THIS IS PROBABLY OFFENDING

#df_overall_baseline_results 
```

df_overall_baseline_results %>%
  mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
  kbl(align = "l", caption = "Sample characteristics at baseline across medication and psychotherapy studies", escape = F) %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>%
  row_spec(row = 1, bold = T) %>%
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>%
  add_indent(c(2, 3, 5, 6, 8, 9)) %>%
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE, results='hide'}

@@

@@

```
### Number of trial sites

There was a significant difference between the number of sites in medication versus psychotherapy trials, as can be seen in Tables 2 and 3. Average number of trial sites was significantly higher across medication studies ( *M* = `r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Medication"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Medication"]`) compared to psychotherapy studies ( *M* =`r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Psychotherapy"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Psychotherapy"]`)(*t* (`r round (sites_t_test$parameter, 2)`) = `r round(sites_t_test$statistic,2)`, *p* =`r sites_t_test$p.value`).

```{r, warning=FALSE, message = FALSE, echo = FALSE, asis = TRUE }

df_sites_summary_stats <- df_sites_summary_stats %>%
  kbl(caption = "Number of sites across medication and psychotherapy studies") %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB") %>%
  column_spec(column = 1, border_right = T)

sites_t_test_stats %>%
  kbl(align = "l", caption = "Results for t-test comparing number of sites between medication and psychotherapy studies") %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB") %>%
  column_spec(column = 1, border_right = T)

@@

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# df_mean_coefs_from_sim

df_mean_coefs_from_sim %>%
  kbl(align = "l", caption = "Summary statistics of the estimated SMDs") %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB") %>%
  column_spec(column = 1, border_right = T)

@@

#aggregate_results_overall
aggregate_results_overall %>%
  kbl(caption = "Meta-regression results") %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB") %>%
  column_spec(column = 1, border_right = T)

@@

group_function <- function(dataframe, my_groups, columns) {
  result_dataframe <- data.frame(
    Group = character(),
    Column = character(),
    N = numeric(),
    Mean = numeric(),
    SD = numeric(),
    Median = numeric(),
    IQR = numeric(),
    stringsAsFactors = FALSE
  )
  unique_groups <- unique(dataframe[[my_groups]])
  for (group in unique_groups) {
    group_data <- dataframe[dataframe[[my_groups]] == group, ]
    for (col in columns) {
      col_data <- group_data[[col]]
      n <- sum(!is.na(col_data))
      mean_val <- mean(col_data, na.rm = TRUE)
      SD <- sd(col_data, na.rm = TRUE)
      median_val <- median(col_data, na.rm = TRUE)
      iqr <- IQR(col_data, na.rm = TRUE)
      result_dataframe <- rbind(result_dataframe,
                                data.frame(Group = group,
                                           Column = col, N = n, Mean = mean_val,
                                           SD = SD, Median = median_val,
                                           IQR = iqr, stringsAsFactors = FALSE))
    }
  }
  return(result_dataframe)
}

@@

control_summary <- control_summary %>% 
  select(-c(Column, Median, IQR))

@@

@@

control_summary %>%
  kbl(align = "l", caption = "Comparing the intensity of the intervention between active and control arms of psychotherapy studies", escape = F) %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  column_spec(1, border_right = T) %>% 
  pack_rows("Number of sessions", 1, 2) %>% 
  pack_rows("Frequency (weeks)", 3, 4) %>% 
  pack_rows("Session length (mins)", 5, 6) %>%
  pack_rows("Total intervention hours", 7, 8) %>%
  pack_rows("Total intervention period (weeks)", 9, 10)

@@

colnames(df_overall_baseline_results)[8] <- paste0(colnames(df_overall_baseline_results)[8], footnote_marker_number(1))

control_summary_t_tests %>%
  kbl(align = "l", caption = "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials") %>% 
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>% 
  row_spec(row = 0, background = "#EBEBEB") %>% 
  column_spec(column = 1, border_right = T)

@@

colnames(df_severity_sens)[8] <- paste0(colnames(df_severity_sens)[8], footnote_marker_number(1))
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
df_severity_sens %>%
  mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
  kbl(align = "l", caption = "Table S1: Sensitivity analyses for comparison of baseline severity levels between psychotherapy and medication trials", escape = F) %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
# Now gender

@@

colnames(df_gender_sens)[8] <- paste0(colnames(df_gender_sens)[8], footnote_marker_number(1))
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
df_gender_sens %>%
  mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
  kbl(align = "l", caption = "Table S2: Sensitivity analyses for comparison of gender composition between psychotherapy and medication trials", escape = F) %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))
```

@@

colnames(df_age_sens)[8] <- paste0(colnames(df_age_sens)[8], footnote_marker_number(1))
```

```{r, warning=FALSE, message = FALSE, echo=FALSE}
df_age_sens %>%
  mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
  kbl(align = "l", caption = "Table S3: Sensitivity analyses for comparison of age between psychotherapy and medication trials", escape = F) %>%
  kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
  row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
  row_spec(row = 1, bold = T) %>% 
  row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
  column_spec(1, border_right = T) %>% 
  add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
  footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))
```

>>>>

---
title: "Fundamental Problems with the Evidence Base of Adolescent Depression Treatments"
author: "Stringaris et al"
prefer-html: true
format: docx
editor: visual
font:
  family: "Calibri"
---

@@

```{r, warning=FALSE, message = FALSE, echo=FALSE}
###########Libraries needed
library(tidyverse)
library(truncnorm)
library(metafor)
library(meta)
library(kableExtra)
library(flextable)
library(magrittr)
library(officer)
```

@@

@@

# Now start looking at age variable

@@

# Rounding
df_overall_baseline_results <- df_overall_baseline_results %>%
  mutate(across(.cols = 2:(ncol(.) - 1), ~ round(., 2))) %>%
  mutate_at(vars(ncol(.)), ~ round(., 3)) %>%
  mutate_all(~if_else(is.na(.), "", as.character(.)))

big_border = fp_border(color="black", width = 1)
small_border = fp_border(color="lightgray", width = 1)

# Create a flextable object
flextable(df_overall_baseline_results) %>%
  set_caption (autonum = TRUE, "Sample characteristics at baseline across medication and psychotherapy studies") %>%
  bold(part = "header") %>%  # Bold the header row
  bold(i = c(1, 4, 7), j = NULL, bold = TRUE, part = "body") %>%  # Bold specific rows
  border_outer(border = small_border) %>%  # Add outer borders
  hline(i = c(1, 2, 4, 5, 7, 8), border = small_border) %>%  # Add light grey horizontal lines all  rows
  hline(i = c(3, 6), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
  hline(part = "header", border = big_border) %>% 
  vline(j = 1, border = big_border) %>% 
  set_table_properties(width = .75) %>%  # Set table width
  bg(bg = "#F2F2F2", part = "header") %>%  # Add light grey background to header row
  autofit() %>%  # Autofit column widths
  font(fontname = "Calibri", part = "all") %>%   # Set font to Calibri 
  padding(i = c(2, 3, 5, 6, 8, 9), j = 1, padding.left = 10) %>% 
  align(j = 2:ncol(df_overall_baseline_results), align = "center", part = "all")

# This doesn't work when rendering to word doc
#   kbl(align = "l", caption = "Sample characteristics at baseline across medication and psychotherapy studies", escape = F) %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>%
#   row_spec(row = 1, bold = T) %>%
#   row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
#   column_spec(1, border_right = T) %>%
#   add_indent(c(2, 3, 5, 6, 8, 9))

@@

### Number of trial sites

@@

colnames(sites_t_test_stats) <- c("Statistic", "df", "p-value", "Lower CI", "Upper CI")

@@

# Create a flextable object

flextable(df_sites_summary_stats) %>%
  set_caption (autonum = TRUE, "Number of sites across medication and psychotherapy studies", style = "Table Caption") %>%
  bold(part = "header") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "all", border = small_border) %>% 
  hline(part = "header", border = big_border) %>% 
  vline(j = 1, border = big_border) %>% 
  set_table_properties(width = .75) %>% 
  bg(bg = "#F2F2F2", part = "header") %>%  
  autofit() %>% 
  font(fontname = "Calibri", part = "all") %>%   
  padding(i = c(1, 2), j = 1, padding.left = 10) %>% 
  align(j = 2:ncol(df_sites_summary_stats), align = "center", part = "all")

flextable(sites_t_test_stats) %>%
  set_caption (autonum = TRUE, "Results for t-test comparing number of sites between medication and psychotherapy studies", style = "Table Caption") %>%
  bold(part = "header") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "header", border = big_border) %>% 
  set_table_properties(width = .75) %>% 
  bg(bg = "#F2F2F2", part = "header") %>%  
  autofit() %>% 
  font(fontname = "Calibri", part = "all") %>%   
  align(align = "center", part = "all")

# This is for if we want to render to pdf

# df_sites_summary_stats <- df_sites_summary_stats %>%
#   kbl(caption = "Number of sites across medication and psychotherapy studies") %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB") %>%
#   column_spec(column = 1, border_right = T)

# sites_t_test_stats %>%
#   kbl(align = "l", caption = "Results for t-test comparing number of sites between medication and psychotherapy studies") %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB") %>%
#   column_spec(column = 1, border_right = T)

```
There was a significant difference between the number of sites in medication versus psychotherapy trials, as can be seen in Tables 2 and 3. Average number of trial sites was significantly higher across medication studies ( *M* = `r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Medication"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Medication"]`) compared to psychotherapy studies ( *M* =`r df_sites_summary_stats$Mean[df_sites_summary_stats$Modality == "Psychotherapy"]`, *SD* =`r df_sites_summary_stats$SD[df_sites_summary_stats$Modality == "Psychotherapy"]`)(*t* (`r round (sites_t_test$parameter, 2)`) = `r round(sites_t_test$statistic,2)`, *p* =`r sites_t_test$p.value`).

@@

flextable(df_mean_coefs_from_sim) %>%
  set_caption (autonum = TRUE, "Summary statistics of the estimated SMDs") %>%
  bold(part = "header") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "header", border = big_border) %>% 
  set_table_properties(width = .75) %>% 
  bg(bg = "#F2F2F2", part = "header") %>%  
  autofit() %>% 
  font(fontname = "Calibri", part = "all") %>%   
  align(j = 2:ncol(df_mean_coefs_from_sim), align = "center")

@@

@@

#aggregate_results_overall
flextable(aggregate_results_overall) %>%
  set_caption (autonum = TRUE, "Metaregression results") %>%
  bold(part = "header") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "header", border = big_border) %>% 
  set_table_properties(width = .75) %>% 
  bg(bg = "#F2F2F2", part = "header") %>%  
  autofit() %>% 
  font(fontname = "Calibri", part = "all") %>%   
  align(j = 2:ncol(aggregate_results_overall), align = "center")

@@

group_function <- function(dataframe, my_groups, columns) {
  result_dataframe <- data.frame(
    Group = character(),
    Column = character(),
    N = numeric(),
    Mean = numeric(),
    SD = numeric(),
    Median = numeric(),
    IQR = numeric(),
    stringsAsFactors = FALSE
  )
  unique_groups <- unique(dataframe[[my_groups]])
  for (group in unique_groups) {
    group_data <- dataframe[dataframe[[my_groups]] == group, ]
    for (col in columns) {
      col_data <- group_data[[col]]
      n <- sum(!is.na(col_data))
      mean_val <- mean(col_data, na.rm = TRUE)
      SD <- sd(col_data, na.rm = TRUE)
      median_val <- median(col_data, na.rm = TRUE)
      iqr <- IQR(col_data, na.rm = TRUE)
      result_dataframe <- rbind(result_dataframe,
                                data.frame(Column = col, Group = group, N = n, Mean = mean_val,
                                           SD = SD, Median = median_val,
                                           IQR = iqr, stringsAsFactors = FALSE))
    }
  }
  return(result_dataframe)
}

@@

control_summary <- control_summary %>% 
  select(-c(Median, IQR))

@@

control_summary <- control_summary %>%
  mutate(Column = case_when(
    Column == "no_sessions" ~ "Number of sessions",
    Column == "freq_weeks" ~ "Frequency (weeks)",
    Column == "length_sessions_mins" ~ "Session length (mins)",
    Column == "total_hours" ~ "Total intervention hours",
    Column == "total_period_weeks" ~ "Total intervention period (weeks)",
    TRUE ~ as.character(Column)  # Keep original value if no match
  ))

@@

as_grouped_data(control_summary, groups = "Column") %>% 
  as_flextable(hide_grouplabel = TRUE ) %>% 
  set_caption (autonum = TRUE, "Comparing the intensity of the intervention between active and control arms of psychotherapy studies") %>%
  bold(part = "header") %>%
  bold(i = c(1, 4, 7, 10, 13), j = NULL, bold = TRUE, part = "body") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "all", border = small_border) %>%  # Add light grey horizontal lines all  rows
  hline(i = c(3, 6, 9, 12), border = big_border) %>%  # Add darker horizontal lines after specific rows %>% 
  hline(part = "header", border = big_border) %>% 
  set_table_properties(width = 1) %>%  
  bg(bg = "#F2F2F2", part = "header") %>%  
  font(fontname = "Calibri", part = "all")

@@

flextable(control_summary_t_tests) %>%
  set_caption (autonum = TRUE, "Results for t-tests comparing intervention intensity between active and control arms of psychotherapy trials", style = "Table Caption") %>%
  bold(part = "header") %>%  
  border_outer(border = small_border) %>%  
  hline(part = "header", border = big_border) %>% 
  set_table_properties(width = .75) %>% 
  bg(bg = "#F2F2F2", part = "header") %>%  
  autofit() %>% 
  font(fontname = "Calibri", part = "all") %>%   
  align(j = 2:ncol(control_summary_t_tests), align = "center", part = "all")

@@

colnames(df_severity_sens)[8] <- paste0(colnames(df_severity_sens)[8], footnote_marker_number(1))

# df_severity_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
#   kbl(align = "l", caption = "Table S1: Sensitivity analyses for comparison of baseline severity levels between psychotherapy and medication trials", escape = F) %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
#   row_spec(row = 1, bold = T) %>% 
#   row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
#   column_spec(1, border_right = T) %>% 
#   add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
#   footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

# Now gender

@@

colnames(df_gender_sens)[8] <- paste0(colnames(df_gender_sens)[8], footnote_marker_number(1))

# df_gender_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
#   kbl(align = "l", caption = "Table S2: Sensitivity analyses for comparison of gender composition between psychotherapy and medication trials", escape = F) %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
#   row_spec(row = 1, bold = T) %>% 
#   row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
#   column_spec(1, border_right = T) %>% 
#   add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
#   footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))

@@

colnames(df_age_sens)[8] <- paste0(colnames(df_age_sens)[8], footnote_marker_number(1))

# df_age_sens %>%
#   mutate_all(~if_else(is.na(.), "", as.character(.))) %>% 
#   kbl(align = "l", caption = "Table S3: Sensitivity analyses for comparison of age between psychotherapy and medication trials", escape = F) %>%
#   kable_material(lightable_options = c("hover"), full_width = FALSE) %>%
#   row_spec(row = 0, background = "#EBEBEB", extra_css = "border-bottom: 1px black;" ) %>% 
#   row_spec(row = 1, bold = T) %>% 
#   row_spec(c(4, 7), bold = T, extra_css = "border-top: 1px solid black;") %>%
#   column_spec(1, border_right = T) %>% 
#   add_indent(c(2, 3, 5, 6, 8, 9)) %>% 
#   footnote(number = c("Statistic derived from the test for subgroup differences in the relevant meta-analysis"))
